Using spatially biased variables in ecosystem condition accounting with a GIS based workflow

### 2024-02-23 manuscript/HIAs.qmd
EAA means and CI

library(tidyverse)
library(knitr)
library(sf)
library(tmap)
libarary(tmaptools)
library(stars)
library(terra)
library(tidyterra)
library(ggtext)
library(cowplot)
library(units)
library(rnaturalearth)
library(rnaturalearthdata)

@@

```{r realValidationPlot}
#| eval: true
# Using a color gradient emphasizes the first color (dark green).
# Lets try discrete colors, and merge some classes to simplify
(validationPlot <- corrCheck |>
  pivot_longer(
    cols = c(i_ADSV, i_alien, i_ditch),
    values_to = "indicatorValue",
    names_to = "indicator",
    values_drop_na = T
  ) |>
  mutate(
    condition = case_when(
      indicatorValue < 0.6 ~ "<0.6",
      indicatorValue < 0.8 ~ "0.6 to 0.8",
      indicatorValue < 0.91 ~ "0.8 to 0.9",
      .default = "0.9 to 1"
    ),
    condition = fct_reorder(condition, indicatorValue),
    indicator = case_when(
      indicator == "i_ADSV" ~ "ADSV",
      indicator == "i_alien" ~ "Alien species",
      indicator == "i_ditch" ~ "Trenching"
    )
  ) |>
  as_tibble() |>
  group_by(indicator, infrastructureIndex, condition) |>
  summarise(n = n()) |>
  ungroup() |>
  group_by(indicator, infrastructureIndex) |>
  mutate(lab = round(n/sum(n)*100),
         lab = case_when(
           lab < 5 ~ NA,
           .default = paste0(lab, "%")
         )) |>
  ggplot(aes(
    x = infrastructureIndex,
    y = n,
    fill = condition
  )) +
  geom_bar(
    position = "fill",
    stat = "identity"
  ) +
  geom_text(aes(label = lab),
    position = position_fill(vjust = 0.5),
    color= "black", vjust = 0.5, size = 4) +
  theme_minimal(base_size = 15) +
  theme(
    panel.grid = element_blank(), 
    axis.text.x = element_text(margin = margin(t = -10)),
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    strip.text = element_textbox(
      size = 12,
      halign = 0.5, linetype = 1, r = unit(5, "pt"), width = unit(1, "npc"),
      padding = margin(2, 0, 1, 0), margin = margin(3, 3, 3, 3))) +
  guides(fill = guide_legend("Indicator values")) +
  xlab("Human Impact Areas") +
  scale_fill_manual(values = c("#E85437","#FBAF00", "#B5DF73", "#009000")) +
  facet_grid(~indicator)
)
```

@@

# Infratructure in each municipality
(infraMuniMap <- tm_shape(muni3) +
  tm_borders() +
  tm_shape(infraMuni3) +
  tm_polygons(
    col = "infrastructureIndex",
    style = "cat",
    title = "Human Impact Areas"
  ) +
  tm_layout(
    legend.show = F,
    panel.label.height = 0) +
  tm_shape(muni3) +
  tm_borders(lwd = 3, col = "black") +
  tm_facets(by = "Municipality"))

# A figure with just the legend
infraMuniMap_l <- tm_shape(muni3) +
  tm_borders() +
  tm_shape(infraMuni3) +
  tm_polygons(
    col = "infrastructureIndex",
    style = "cat",
    title = "Human\nImpact\nArea"
  ) +
  tm_layout(legend.only = TRUE,
            legend.position = c("left", "bottom"),
            legend.outside = F)

@@

# tmap_save(methodsMap, "../figures/studyLocations.tiff",
#           dpi= 1000,
#           units = "cm",
#           width = 18,
#           height = 10)
#
# saveRDS(methodsMap, "../figures/studyLocation.RDS")
```

@@

```{r test-function}
#| eval: false
test_sd <- wgt_mean(x = nature3$i_ADSV[!is.na(nature3$i_ADSV)], 
                 weights = drop_units(nature3$area[!is.na(nature3$i_ADSV)]), 
                 sigma.x = national_sd_ADSV,
                 stat = "sd")
test_mean <- wgt_mean(x = nature3$i_ADSV[!is.na(nature3$i_ADSV)], 
                 weights = drop_units(nature3$area[!is.na(nature3$i_ADSV)]), 
                 sigma.x = national_sd_ADSV,
                 stat = "mean")

hist(rnorm(1000, test_mean, test_sd))

@@

stats_tbl |>
  mutate(
    indicator = case_when(
      indicator == "i_ADSV" ~ "ADSV",
      indicator == "i_alien" ~ "Alien species",
      indicator == "i_ditch" ~ "Trenching"
    )
  ) |>
  rowwise() |>
  mutate(
    low = quantile(rnorm(200, mean, sd), probs = 0.025),
    high = quantile(rnorm(200, mean, sd), probs = 0.975),
    high = ifelse(high > 1, 1, high)
  ) |>
  ggplot(aes(x=infrastructureIndex, 
             y=mean, 
             ymin=low,
             ymax=high,
             col=factor(infrastructureIndex),
             fill=factor(infrastructureIndex))) + 
  geom_linerange(
    size=10,
  geom_point(
    size=3, 
    shape=21, 
    colour="white", 
    stroke = 0.5,
    ) +
  geom_text(
    aes(y = low, label = n),
    nudge_y = -0.1,
    show.legend = F
  ) +
  scale_fill_manual(values=dotCOLS)+
  scale_color_manual(values=barCOLS)+
  scale_x_discrete(name="") +
  scale_y_continuous(name="Indicator values", limits = c(-0.1, 1)) +
  coord_flip() +
  theme_bw() +
  labs(fill = "HIA",
       col = "HIA") +
  facet_grid(indicator ~ Municipality)
```

```{r}
# Next I need to 
#  - copy the weighted means and SDs over to the ecosystem delineation map polygons 
#  - show intermittent resulting map with ggmagnify
#  - sample from these distributions (ignore missing HIAs)
```

```{r}

@@

@@

In this study we explore the potential for using a stratified aggregation technique to make use of spatially biased field data in ecosystem condition accounting.
We demonstrate this technique using a generic GIS-based workflow for compiling ecosystem condition accounts that can be applied at any spatial scale, and we highlight the opportunities for local use-cases of this workflow by contrasting our findings across four neighboring municipalities in Norway.
The main question is how much generalisation can we perform on the data we have before the resulting indicator looses its practical value in local governance processes.
We end by interviewing end-users from the relevant municipalities about the perceived benefits and shortcomings of our condition indicators.

@@

We started by defining the reference state as one where ecosystems are subject to little or no human influence, with a climate as in the period 1961-1990 and a native species pool similar as today (i.e no mammoths allowed).
We then identified two *ecosystem condition characteristics* that, along with many other characteristics that we do not specify here, define the typical behavior of open mire ecosystems in the reference condition: species composition, and; vegetation intactness.

To describe these two characteristics we use six condition variables from the same data set ( @tbl-variables).

See @miljødirektoratet for the sampling protocol (in Norwegian).

| id  | Variable code        | Variable name                                    | Measurement unit                                                                        | Description                                                                                                                                                       | Reference                          |
|----------|----------|----------|----------|-----------------------|----------|
| 1   | 7FK                  | Prevalence of alien species                      | Unitless, ordinal, non-linear scale from 1 (no alien species) to 7 (only alien species) | The fraction of the species composition made up from alien species                                                                                                | @runehalvorsen2019                 |
| 2   | 7SE                  | Human caused abration or abration-caused erosion | Unitless, ordinal, non-linear scale from 1 to 4.                                        | Measures the frequency of imagined 4 m^2^ quadrats layed over the area that has some sign of abration                                                             | @runehalvorsen2019                 |
| 3   | PRSL                 | *as above*                                       | Unitless, ordinal, non-linear scale from 0 to 7.                                        | Same as 7SE, but recorded at a higher resolution                                                                                                                  | @miljødirektoratet                 |
| 4   | 7TK                  | Tracks from large vehicles                       | Unitless, ordinal, non-linear scale from 1 to 4.                                        | Measures the frequency of imagined 100 m^2^ quadrats layed over the area that has some signs of vehicle tracks                                                    | @rune_halvorsen_dokumentasjon_2019 |
| 5   | PRTK                 | *as above*                                       | Unitless, ordinal, non-linear scale from 0 to 7.                                        | Same as 7TK, but recorded at a higher resolution                                                                                                                  | @miljødirektoratet                 |
| 6   | 7GR-GI               | Ditching intensity                               | Unitless, ordinal scale from 1 to 5                                                     | Describes the effect that drainage ditches is estimate to have on the species composition and environmental variables ones the system reached its new equilibrium | @miljødirektoratet                 |
| 7   | Infrastructure Index | Infrastructure Index                             | Unitelss linear scale from 0 to 13.2                                                    | Unitelss index ranging from from 0 to 13.2                                                                                                                        | @erikstad_index_2023               |
| 8   | HIA                  | Human Impact area                                | Ordinal, non-linear scale from 1 to 4                                                   | A categorical representation of the infrastructure index                                                                                                          | this paper                         |

@@

Variables 1 -5 were 7FK were originally recorded along frequency ranges.
We the lower limit for each frequency range to convert them into percentages.
This was done because the data was strongly right skewed .
7FK was developed into a single indicator (i.e. a normalised variable) named *Alien plant cover*.
Variables 2-5 describe very related aspects and so they where combined into one indicator called antropogenic disturbance to soil and vegetation, or ADSV for short.
This was done by summing the two variables after they had been converted to percentages.
This was not a perfect solution, especially since some localities only had one of the variables recorded, but we chose this, rather than for example using a *worst rule* principle, in order to better separate the localities in terms of their indicator values.
For both indicators (alien species and ADSV), the lower and upper reference values, i.e. the worst and best possible condition that the variables can be in, is defined as 100% and 0%, respectively.
The threshold for *good ecosystem condition* was defined as 10%, and this was mapped to the value 0.6 on the rescaled indicators.
Variable 7GR-GI was rescaled into an indicator named *ditching intensity* by using the lower and upper reference values 1 and 5, respectively, and a threshold value of 2.5.
A variable value of 1 indicates an intact mire, and a value of 5 indicates a mire transitioning away from a wetland.
A value of 2 indicator observable change within the range expected for the same mapping unit, and a value of 3 indicates a mire transitioning into a neighboring (ecologically speaking) mapping unit.

The .
See @fig-workflow for a schematic workflow for the indicator development.

@@

```{r fig-studyLocation}
#| eval: true
#| include: true
#| out.width: '100%'
#| fig.cap:
#| - "Postion of the three example municipalities in
#| Norway. For each municipality the map shows ocean in blue and non-ocean
#| in green. The survey coverage maps are in grey, and the mapped mires are
#| in red, with polygon borderes made extra thich to make them visible,
#| but then also exagerating their size."

#(temp <- readRDS("../figures/studyLocation.RDS"))
positionMap
```

```{r fig-studyLocation2}
#| eval: true
#| include: true
#| out.width: '100%'
#| fig.cap:
#| - "Map of the three focal municiaplities. For each municipality the map shows 
#| ocean in blue and non-ocean
#| in green. The survey coverage maps are in grey, and the mapped mires are
#| in red, with polygon borderes made extra thich to make them visible,
#| but then also exagerating their size."

#(temp <- readRDS("../figures/studyLocation.RDS"))
methodsMap
```

@@

```{r fig-validation}
#| eval: true
#| include: true
#| out.width: '100%'
#| fig.cap:
#| - "Proportion of localities with different scaled indicator values for three 
#| ecosystem condition indicators along a Human Impact Gradient"

validationPlot
```

National/regional indicator - do we have enough data?

Municipal case study

@@

# References {.unnumbered}

>>>>

library(tidyverse)
library(knitr)
library(sf)
library(tmap)
library(tmaptools)
library(stars)
library(terra)
library(tidyterra)
library(ggtext)
library(cowplot)
library(units)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggmagnify)

@@

```{r realValidationPlot}
#| eval: true
# Using a color gradient emphasizes the first color (dark green).
# Lets try discrete colors, and merge some classes to simplify
(validationPlot <- corrCheck |>
  pivot_longer(
    cols = c(i_ADSV, i_alien, i_ditch),
    values_to = "indicatorValue",
    names_to = "indicator",
    values_drop_na = T
  ) |>
  mutate(
    condition = case_when(
      indicatorValue < 0.6 ~ "<0.6",
      indicatorValue < 0.8 ~ "0.6 to 0.8",
      indicatorValue < 0.91 ~ "0.8 to 0.9",
      .default = "0.9 to 1"
    ),
    condition = fct_reorder(condition, indicatorValue),
    indicator = case_when(
      indicator == "i_ADSV" ~ "ADSV",
      indicator == "i_alien" ~ "Alien species",
      indicator == "i_ditch" ~ "Trenching"
    )
  ) |>
  as_tibble() |>
  group_by(indicator, infrastructureIndex, condition) |>
  summarise(n = n()) |>
  ungroup() |>
  group_by(indicator, infrastructureIndex) |>
  mutate(lab = round(n/sum(n)*100),
         lab = case_when(
           lab < 5 ~ NA,
           .default = paste0(lab, "%")
         )) |>
  ggplot(aes(
    x = infrastructureIndex,
    y = n,
    fill = condition
  )) +
  geom_bar(
    position = "fill",
    stat = "identity"
  ) +
  geom_text(aes(label = lab),
    position = position_fill(vjust = 0.5),
    color= "black", vjust = 0.5, size = 4) +
  theme_minimal(base_size = 15) +
  theme(
    panel.grid = element_blank(), 
    axis.text.x = element_text(margin = margin(t = -10)),
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    strip.text = element_textbox(
      size = 12,
      halign = 0.5, linetype = 1, r = unit(5, "pt"), width = unit(1, "npc"),
      padding = margin(2, 0, 1, 0), margin = margin(3, 3, 3, 3))) +
  guides(fill = guide_legend("Indicator values")) +
  xlab("Homogeneous Impact Areas") +
  scale_fill_manual(values = c("#E85437","#FBAF00", "#B5DF73", "#009000")) +
  facet_grid(~indicator)
)

ggsave("../images/validation-plot.jpg",
       plot=validationPlot,
       width = 9,
  height = 5)
```

@@

# Infratructure in each municipality
(infraMuniMap <- tm_shape(muni3) +
  tm_borders() +
  tm_shape(infraMuni3) +
  tm_polygons(
    col = "infrastructureIndex",
    style = "cat",
    title = "Homogeneous Impact Areas"
  ) +
  tm_layout(
    legend.show = F,
    panel.label.height = 0) +
  tm_shape(muni3) +
  tm_borders(lwd = 3, col = "black") +
  tm_facets(by = "Municipality"))

# A figure with just the legend
infraMuniMap_l <- tm_shape(muni3) +
  tm_borders() +
  tm_shape(infraMuni3) +
  tm_polygons(
    col = "infrastructureIndex",
    style = "cat",
    title = "Homogeneous\nImpact Area"
  ) +
  tm_layout(legend.only = TRUE,
            legend.position = c("left", "bottom"),
            legend.outside = F)

@@

#tmap_save(methodsMap, "../images/studyLocations.tiff",
#           dpi= 1000,
#           units = "cm",
#           width = 18,
#           height = 10)
#tmap_save(methodsMap, "../images/studyLocations.jpg",
#           units = "cm",
#           width = 18,
#           height = 10)

#
#saveRDS(methodsMap, "../figures/studyLocation.RDS")
```

@@

@@

(forest_plot <- stats_tbl |>
  mutate(
    indicator = case_when(
      indicator == "i_ADSV" ~ "ADSV",
      indicator == "i_alien" ~ "Alien species",
      indicator == "i_ditch" ~ "Trenching"
    )
  ) |>
  rowwise() |>
  mutate(
    low = quantile(rnorm(200, mean, sd), probs = 0.025),
    high = quantile(rnorm(200, mean, sd), probs = 0.975),
    high = ifelse(high > 1, 1, high)
  ) |>
  ggplot(aes(x=infrastructureIndex, 
             y=mean, 
             ymin=low,
             ymax=high,
             col=factor(infrastructureIndex),
             fill=factor(infrastructureIndex))) + 
  geom_linerange(
    size=10) +
  geom_point(
    size=3, 
    shape=21, 
    colour="white", 
    stroke = 0.5,
    ) +
  geom_text(
    aes(y = low, label = n),
    nudge_y = -0.1,
    show.legend = F
  ) +
  scale_fill_manual(values=dotCOLS)+
  scale_color_manual(values=barCOLS)+
  scale_x_discrete(name="") +
  scale_y_continuous(name="Indicator values", limits = c(-0.1, 1)) +
  coord_flip() +
  theme_bw() +
  labs(fill = "HIA",
       col = "HIA") +
  facet_grid(indicator ~ Municipality)
)

#ggsave("../images/forest-plot.jpg",
#       plot=forest_plot)
```

```{r}
# Next I need to 
#  - copy the weighted means and SDs over to  HIA map and then over to the 
#    ecosystem delineation map polygons. 
#  - show intermittent resulting map with ggmagnify
#  - sample from these distributions (ignore missing HIAs)
```

```{r}
#| eval: true
#| cache: true

# Copy the weighted means and SDs over to HIA map and then over to the 
# ecosystem delineation map polygons. 
spread_na <- mire_stars_na |>
  st_as_sf(merge=T) |>
  filter(Myr153 == 1) |>
  st_intersection(infraMuni3 |> select(infrastructureIndex)) |>
  mutate(area = geometry |> st_area()) |>
  left_join(stats_tbl |> 
              as_tibble() |>
              filter(Municipality == "Nord-Aurdal") |>
              select(mean, sd, indicator, n, infrastructureIndex, Municipality),
            by = "infrastructureIndex")

spread_nf <- mire_stars_nf |>
  st_as_sf(merge=T) |>
  filter(Myr153 == 1) |>
  st_intersection(infraMuni3 |> select(infrastructureIndex)) |>
  mutate(area = geometry |> st_area()) |>
  left_join(stats_tbl |> 
              as_tibble() |>
              filter(Municipality == "Nordre Follo") |>
              select(mean, sd, indicator, n, infrastructureIndex, Municipality),
            by = "infrastructureIndex")

spread_gr <- mire_stars_gr |>
  st_as_sf(merge=T) |>
  filter(Myr153 == 1) |>
  st_intersection(infraMuni3 |> select(infrastructureIndex)) |>
  mutate(area = geometry |> st_area()) |>
  left_join(stats_tbl |> 
              as_tibble() |>
              filter(Municipality == "Gran") |>
              select(mean, sd, indicator, n, infrastructureIndex, Municipality),
            by = "infrastructureIndex")

@@

```{r example-spread}
#| eval: true
#| cache: true

# A plot of Nord-Aurdal with the mean indicator values per HIA spread over the 
# EDM
from <- c(xmin = 510000, xmax = 515000, ymin = 6741000, ymax = 6746000)
to <-   c(xmin = 495000, xmax = 520000, ymin = 6765000, ymax = 6790000)
myCols <- c(
  #"#E85437",
  "#FBAF00", 
  "#B5DF73", 
  "#009000"
  )

spread_na_map <- spread_na |>
  filter(indicator == "i_ditch") |>
  mutate(Trenching = factor(round(mean, 2))) |>
  ggplot() +
  geom_sf(aes(fill=Trenching,
              color = Trenching)) +
  geom_sf(data = na,
          alpha=0) +
  scale_fill_manual(values = myCols) +
  scale_color_manual(values = myCols) +
  coord_sf(
    datum = st_crs(myCRS),
    xlim = c(494174.8 , 537114.7 ), 
    ylim = c(6737092 , 6789676)) +
  ggmagnify::geom_magnify(from = from, to = to, 
                          expand = 0,
                          shadow =T,
                          corners = 0.1) +
  theme_bw()

ggsave("../images/spread-na.tiff",
       plot = spread_na_map)
ggsave("../images/spread-na-small.tiff",
       plot = spread_na_map,
       dpi=150)
ggsave("../images/spread-na-small.jpg",
       plot = spread_na_map)
```

```{r EAA}
#| eval: true
#| cache: true

combineALl <- rbind(
  spread_nf,
  spread_gr,
  spread_na
) |>
  as_tibble() |>
  drop_na() |>
  group_by(Municipality, indicator) |>
  rowwise() |>
  mutate(
    i_sample = list(sample(rnorm(10, mean, sd)))) |>
  unnest(cols=i_sample) |>
  summarise(mean = mean(i_sample),
            lower = quantile(i_sample, probs = c(0.025)),
            upper = quantile(i_sample, probs = c(0.975)))

# add area weights to the above and make forest plot

#combineALl
#tail(combineALl)
#combineALl$i_sample[2]
```

@@

In this study we explore the potential for using a stratified aggregation technique to make use of spatially biased field data in ecosystem condition accounting.
We demonstrate this technique using a generic GIS-based workflow for compiling ecosystem condition accounts that can be applied at any spatial scale, and we highlight the opportunities for local use-cases of this workflow by contrasting our findings across three neighboring municipalities in Norway (@fig-studyLocation).
The main question is how much generalisation can we perform on the data we have before the resulting indicator looses its practical value in local governance processes.
We end by interviewing end-users from the relevant municipalities about the perceived benefits and shortcomings of our condition indicators.

```{r fig-studyLocation}
#| eval: true
#| include: true
#| out.width: '100%'
#| fig.cap:
#| - "Postion of the three example municipalities in
#| Norway. "

positionMap
```

@@

We started by defining the reference state as one where ecosystems are subject to little or no human influence, with a climate as in the period 1961-1990 and a native species pool similar as today (i.e no mammoths allowed).
We then identified two *ecosystem condition characteristics* that, along with many other characteristics that we do not specify here, define the typical behavior of open mire ecosystems in the reference condition: species composition, and; vegetation intactness.
To describe these two characteristics we use six condition variables from the same data set ( @tbl-variables).
See @miljødirektoratet for the sampling protocol (in Norwegian).

```{r fig-studyLocation2}
#| eval: true
#| include: true
#| out.width: '80%'
#| fig.cap:
#| - "Map of the three focal municiaplities. For each municipality the map shows 
#| ocean in blue and non-ocean
#| in green. The survey coverage maps are in grey, and the mapped mires are
#| in red, with polygon borderes made extra thich to make them visible,
#| but then also exagerating their size."

knitr::include_graphics("../images/studyLocations.jpg")
```

<!--# Three characteristics now? -->

| id  | Variable code        | Variable name                                    | Measurement unit                                                                        | Description                                                                                                                                                       | Reference                          |
|-------|---------|-----------|--------------|----------------------|----------|
| 1   | 7FK                  | Prevalence of alien species                      | Unitless, ordinal, non-linear scale from 1 (no alien species) to 7 (only alien species) | The fraction of the species composition made up from alien species                                                                                                | @runehalvorsen2019                 |
| 2   | 7SE                  | Human caused abration or abration-caused erosion | Unitless, ordinal, non-linear scale from 1 to 4.                                        | Measures the frequency of imagined 4 m^2^ quadrats layed over the area that has some sign of abration                                                             | @runehalvorsen2019                 |
| 3   | PRSL                 | *as above*                                       | Unitless, ordinal, non-linear scale from 0 to 7.                                        | Same as 7SE, but recorded at a higher resolution                                                                                                                  | @miljødirektoratet                 |
| 4   | 7TK                  | Tracks from large vehicles                       | Unitless, ordinal, non-linear scale from 1 to 4.                                        | Measures the frequency of imagined 100 m^2^ quadrats layed over the area that has some signs of vehicle tracks                                                    | @rune_halvorsen_dokumentasjon_2019 |
| 5   | PRTK                 | *as above*                                       | Unitless, ordinal, non-linear scale from 0 to 7.                                        | Same as 7TK, but recorded at a higher resolution                                                                                                                  | @miljødirektoratet                 |
| 6   | 7GR-GI               | Ditching intensity                               | Unitless, ordinal scale from 1 to 5                                                     | Describes the effect that drainage ditches is estimate to have on the species composition and environmental variables ones the system reached its new equilibrium | @miljødirektoratet                 |
| 7   | Infrastructure Index | Infrastructure Index                             | Unitelss linear scale from 0 to 13.2                                                    | Unitelss index ranging from from 0 to 13.2                                                                                                                        | @erikstad_index_2023               |
| 8   | HIA                  | Human Impact area                                | Ordinal, non-linear scale from 1 to 4                                                   | A categorical representation of the infrastructure index                                                                                                          | this paper                         |

@@

Variables 1 -5 were 7FK were originally recorded along frequency ranges.
We the lower limit for each frequency range to convert them into percentages.
This was done because the data was strongly right skewed .
7FK was developed into a single indicator (i.e. a normalised variable) named *Alien plant cover*.
Variables 2-5 describe very related aspects and so they where combined into one indicator called antropogenic disturbance to soil and vegetation, or ADSV for short.
This was done by summing the two variables after they had been converted to percentages.
This was not a perfect solution, especially since some localities only had one of the variables recorded, but we chose this, rather than for example using a *worst rule* principle, in order to better separate the localities in terms of their indicator values.
For both indicators (alien species and ADSV), the lower and upper reference values, i.e. the worst and best possible condition that the variables can be in, is defined as 100% and 0%, respectively.
The threshold for *good ecosystem condition* was defined as 10%, and this was mapped to the value 0.6 on the rescaled indicators.
Variable 7GR-GI was rescaled into an indicator named *ditching intensity* by using the lower and upper reference values 1 and 5, respectively, and a threshold value of 2.5.
A variable value of 1 indicates an intact mire, and a value of 5 indicates a mire transitioning away from a wetland.
A value of 2 indicator observable change within the range expected for the same mapping unit, and a value of 3 indicates a mire transitioning into a neighboring (ecologically speaking) mapping unit.
See @fig-workflow for a schematic workflow for the indicator development.

@@

{{< pagebreak >}}

@@

```{r fig-validation}
#| eval: true
#| include: true
#| out.width: '90%'
#| fig.cap:
#| - "Proportion of localities with different scaled indicator values for three 
#| ecosystem condition indicators along a gradient of infrastructure
#| densities."

knitr::include_graphics("../images/validation-plot.jpg")
```

```{r fig-forest}
#| eval: true
#| include: true
#| out.width: '80%'
#| fig.cap:
#| - "Indicator values (circles = mean; bars = 95% credible intervals) for three
#| mire ecosystem condition indicators in three municipalities. The indicator values 
#| are calculated uniquely for each Homogeneous Impact Area (HIA) in each 
#| municipality. The numbers to the left of each bar is the sample size, i.e. 
#| the number of surveyed mires."

knitr::include_graphics("../images/forest-plot.jpg")
```

```{r fig-spread-example}
#| eval: true
#| include: true
#| out.width: '80%'
#| fig.cap:
#| - "An indicator for mire trenching shown for Nord-Aurdal municipality. 
#| Individual mire polygons are colored by the mean indicator values for the
#| homogenous impact area where it lies. Colors are chosen to best reflect 
#| categorical differences and exagerates the absolute difference between 
#| areas. The inset it just a visual aid."

knitr::include_graphics("../images/spread-na-small.jpg")
```

@@

{{< pagebreak >}} \# References {.unnumbered}
